### 2016-12-07 paper/chapters/more.tex
added GAN architecture image

\section{More Generative Directed Models}
\label{sec:more}

\subsection{Generative Moment Matching Networks}
\label{sub:more_mmn}

\subsection{Auto-Regressive Networks}
\label{sub:more_arn}

%See \cite{gan_openai:2016} and \cite{gan_training:2016}.