### 2016-12-07 paper/chapters/vae/cvae.tex
added some text to cvae and iwae

Instead of randomly generated data, generation of output conditioned on some prior information is needed.
A few examples would be super-resolution of images (citation!), prediction of subsequent frames of an video or other prediction tasks

Conditional VAE is a modification of the original VAE framework to support conditioning on another variable $y$ in addition to the input data $x$.

To support conditional distributions, the following relations have to be rewritten:
\begin{itemize}
\item \textbf{Sampling from data space}: $x \sim p_\theta(x|y,z)$ instead of $x \sim p_\theta(x|z)$.\\
\item \textbf{Probabilistic encoder}: $q_\theta(z|x,y)$ instead of just $q_\theta(z|x)$.\\
\item \textbf{Probabilistic decoder}: $p_\phi(x|y,z)$ instead of just $p_\phi(x|z)$.\\
\end{itemize}