### 2016-12-08 paper/media/vae_conceptual.tex
added conceptual vae figure

\documentclass{standalone}
\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
\usetikzlibrary{positioning, arrows.meta, calc, shapes.misc}

\begin{document}

\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}

\tikzstyle{gb} = [draw,outer sep=3,inner sep=3,minimum size=10,line width=1, very thick, draw=black!55, top color=white,bottom color=black!20]
\tikzstyle{arr} = [thick,-{Latex[length=3mm,width=2mm]}]
\tikzstyle{cloud} = [draw, circle, node distance=2.5cm, minimum height=2em, gb]

\begin{tikzpicture}

\begin{scope}[xshift=5cm,on grid]

% input space
\node[gb] (input) at (0,0) {$x$};
\node[fill=black!20] (lbl_input) at (0,1) {Input};

% output / reconstruction
\node[gb] (output) at ($(input) + (5.5cm,0)$) {$\tilde x$};
\node[fill=black!20] (lbl_rec) at ($(output) + (0,1cm)$) {Reconstruction};

% latent space
\node[fill=black!20] (lbl_lat) at ($(input) + (2.7cm,1cm)$) {Latent space};

% draw plot for normal dist
\begin{axis}[
  no markers,
  domain=0:6,
  samples=100,
  ytick=\empty,
  xtick=\empty,
  xlabel={$(\mu,\sigma)$},
  width=3cm,
  at={($(input) + (2cm,-0.5cm)$)}
  ]
  \addplot [very thick,cyan!50!black, domain=0:6] {gauss(3,1)};
\end{axis}

% input -> latent
\draw[arr] (input) -- node[label=above:$q_\phi(z|x)$] {}($(input) + (2cm,0)$) {};

% latent -> output (reconstruction)
\draw[arr] ($(input) + (3.5cm,0)$) -- node[label=above:$p_\theta(x|z)$] {}(output) {};

\end{scope}

\end{tikzpicture}
\end{document}