### 2016-12-12 paper/chapters/vae/draw.tex
sbn + VI + draw

The \emph{deep recurrent attention writer}, in short DRAW is an advanced model uses the selective attention mechanism in addition to a VAE with both recurrent encoder and recurrent decoder.
DRAW allows to basically the pixels of an image sequentially very human-like.