### 2016-12-12 paper/chapters/vae/iwae.tex

% write about multiple stochastic layers
Until now, the presented model includes only one stochastic layer.
\cite{dlgm:2014} have presented a model for learning through multiple stochastic layers which can be easily applied to the VAE framework as shown in \ref{fig:vae_multiple_layers}.

\begin{figure}[htb]
\centering
\includestandalone[width=\linewidth,mode=buildnew]{media/vae_multiple_layers}
  \caption{VAE with multiple stochastic layers }\label{fig:vae_multiple_layers}
  \medskip
  \small
  The left upward arrows indicate the recognition networks $q(h_1|x)$ and $q(h_2|h_1)$.
  Note that the recognition networks can have multiple deterministic layers (including non-linearities).\\

  $p(h_2)$ denotes the probability distribution over the highest(?) latent space.\\

  On the right, the two conditional distributions $p(h_1|h_2)$ and $p(x|h_2)$ are representing two networks used for the generative process.
\end{figure}
% write about k passes (just averages)

% write about IWAE, extends these k passes with importance weights
Importance Weighted Autoencoders (IWAE) extends on the VAE framework using $k$-sample importance weighting of the log-likelihood.
% add intuition for IWAE, ruslan explained it very well during his DL summer school talk! (talk_Montreal_2016_Salakhutdinov.pdf)
Using importance weighting can be informally interpreted as putting more weight on data with high likelihood.

The authors of the papers argue that using more samples will only tighten the lower variational bound.
$$
\log p(x) \geq \mathcal{L}_{k+1}(x) \geq \mathcal{L}_{k}(x)
$$

Additionally, if $\frac{p(h,x)}{q(h|x)}$ is bounded, then the model is guaranteed to reach the optimimum:
$$
\lim_{k \rightarrow \infty} \mathcal{L}_{k}(x) = \log p(x)
$$

Of course, using more samples will result in higher computational cost.