### 2016-12-18 paper/chapters/overview.tex
VAE & GAN -> differentiable generator nets

\section{Overview of Directed Generative Models}
\label{sec:overview}

\paragraph{Sigmoid Belief Nets (SBN)}
\label{par:overview_sbn}
SBNs are one of the earliest generative models, using a simple architecture in which sampling from is easy, but learning and performing inference is not.

\paragraph{Differentiable Generator Networks}
Due to the success of gradient-based methods for learning deep networks, a large part of recent models are using a differentiable generator network which allows for error backpropagation.\\
Most prominent examples of models with differentiable generator nets are the variational autoencoder (VAE) \cite{vae:2014} and generative adversarial networks (GAN) \cite{gan:2014}.
These models include a generator networks representing a family of parametrized non-linear functions $g$ implemented usually implemented as neural networks.\\
Using $g$ as distribution transformation allows samples $z$ to be transformed into $x$ where the distribution over $x$ may be more complicated than $z$.
The parameters of $g$ can be inferred using training data and an appropiate loss function.\\
However, instead of maximizing the log-likelihood of $p(x)$ directly we choose $g$ to emit not samples of $x$ but parameters of a conditional distribution over $x$.

\paragraph{Auto-regressive networks}

%\paragraph{Variational Autoencoders (VAE) \cite{vae:2013}\cite{rezende:2014}}
%\label{par:overview_vae}
%Variational Autoencoders (VAE) use two networks which are trained jointly using approximated learned inference.
%The objective function is aimed at building a reasonable simple latent space and approximating the intractable posterior distribution using variational inference.
%It consists of the recognition and inference network which can be understood in the autoencoder jargon as probabilistic encoder and decoder.
%Using variational bayesian methods the intractable posterior distribution can be approximated and learned using backpropagation and gradient-based methods.

%But in addition to regular autoencoders encourages the VAE the latent space to match a specific probability distribution.
%This encouragement is done using a variational lower bound on the distribution output by the encoder network.

%where the encoder takes input data and translates it into a latent structure while
%the decoder takes the latent structure and produces data which is as close as possible to the initial input data.

%In the case of VAEs, the distribution is intractable and we can only approximate the posterior distribution.
% intractable -> use neural network, which we can train
% --> no closed form possible
%We use the KL-Divergence to train the network.

% TODO read chapter 19, approximate inference
%In VAEs however, this

%\paragraph{Generative Adversarial Networks\cite{gan:2014}}
%\label{par:overview_gan}
%GANs are a game theoretic approach to generative modelling
%in which two networks compete against each other.
%The discriminator $D$ is trying to distinguish the generated samples
%by the generator $G$ from real world data.
%The goal for $G$ however is to produce data as realistic as possible.\\
%During the early phase of training, $G$ produces random noise which
%is easy to differentiate from groundtruth data.
%Both networks are trained in parallel, where both networks try to get better
%at their respective objective.
%In the best-case scenario is $G$ able to produce data indistinguishly from real world data in which
%case the discriminator will only be able to guess correctly with a 50\% chance.
%This plateau is called Nash equilibrium (citation needed).