### 2017-01-06 paper/chapters/more.tex

\newpage
\section{More Generative Directed Models}
\label{sec:more}

There are several more deep directed generative models than shown here, however in order to discuss some models more in detail only two recent promising models have been presented.

\subsection{Generative Moment Matching Networks}
\label{sub:more_mmn}
Generative moment matching networks (GMMN) present a similar idea to GANs, however instead of posing the generative modelling problem as game it is optimized using \emph{maximum mean discrepancy} (MMD).

\vspace{10cm}

\subsection{Auto-Regressive Networks}
\label{sub:more_arn}

%See \cite{gan_openai:2016} and \cite{gan_training:2016}.