### 2017-01-09 slides/figures/vae_conceptual.tex
synced

\documentclass{standalone}

\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
\usetikzlibrary{fit,backgrounds}

\usetikzlibrary{positioning, arrows.meta, calc, shapes.misc}

\begin{document}

\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}

\tikzstyle{gb} = [draw,outer sep=0.1cm,inner sep=0.3cm,minimum size=10,line width=0.5, draw=black!55, top color=white,bottom color=black!20]
\tikzstyle{arr} = [thick,-{Latex[length=3mm,width=2mm]}]
\tikzstyle{cloud} = [draw, circle, node distance=2.5cm, minimum height=2em, gb]

\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\begin{tikzpicture}

%\begin{scope}[xshift=5cm,on grid]

% input space
\node[gb] (input) at (0cm,0cm) {$x$};
\node[fill=black!20] (lbl_input) at (0,-1) {Input};

% output / reconstruction
\node[gb] (output) at ($(input) + (3.0cm,0.0cm)$) {$\tilde x$};
\node[fill=black!20] (lbl_rec) at ($(output) + (0,-1cm)$) {Reconstruction};

% latent space
\node[fill=black!20] (lbl_lat) at ($(input) + (2.7cm,5.2cm)$) {$z \sim \mathcal{N}(\mu, \sigma^2 \odot I)$};

    %\begin{pgfonlayer}{background}

% draw plot for normal dist
\begin{axis}[
  no markers,
  domain=0:6,
  samples=100,
  ytick=\empty,
  width=3.5cm,
  xtick=\empty,
  xlabel={$(\mu,\sigma)$},
  at={($(input) + (-0.95cm,4.5cm)$)}
  ]
  \addplot [thick,cyan!50!black, domain=0:6] {gauss(3,1)};
  \node[] (left_bot) at (axis cs:0.0,-0.15) {};
  \node[] (left_top_aux) at (axis cs:0.0,0.2) {};
\end{axis}

  \node [above right=of left_top_aux] (left_top) {Latent space};

    %\end{pgfonlayer}
\begin{scope}[on background layer]
  \node (outer) [dashed,draw,fill=blue!10,fit=(left_bot) (left_top) (lbl_lat)] {};
\end{scope}

% input -> latent
\draw[arr] (input) -- node[label=left:$q_\phi(z|x)$] {}($(input) + (0.0cm,3.8cm)$) {};

% latent -> output (reconstruction)
\draw[arr] ($(input) + (3.0cm,4.5cm)$) -- node[label=right:$p_\theta(x|z)$] {}(output) {};

%\end{scope}

\end{tikzpicture}
\end{document}