### 2017-01-09 slides/figures/vae_conceptual.tex.preview.tex

\documentclass{standalone}
\usepackage{pgfplots}
\usetikzlibrary{fit,backgrounds}

\pgfplotsset{compat=1.14}
\usetikzlibrary{positioning, arrows.meta, calc, shapes.misc}

\begin{document}

\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}

\tikzstyle{gb} = [draw,outer sep=0.1cm,inner sep=0.3cm,minimum size=10,line width=0.5, draw=black!55, top color=white,bottom color=black!20]
\tikzstyle{arr} = [thick,-{Latex[length=3mm,width=2mm]}]
\tikzstyle{cloud} = [draw, circle, node distance=2.5cm, minimum height=2em, gb]

\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\begin{tikzpicture}

%\begin{scope}[xshift=5cm,on grid]

% input space
\node[gb] (input) at (0cm,0cm) {$x$};
\node[fill=black!20] (lbl_input) at (0,-1) {Input};

% output / reconstruction
\node[gb] (output) at ($(input) + (3.0cm,0.0cm)$) {$\tilde x$};
%\node[fill=black!20] (lbl_rec) at ($(output) + (0,1cm)$) {Reconstruction};

% latent space
\node[fill=black!20] (lbl_lat) at ($(input) + (2.7cm,5.2cm)$) {$z \sim \mathcal{N}(\mu, \sigma^2 \odot I)$};

%\begin{scope}
    %\begin{pgfonlayer}{background}
%\begin{scope}[on background layer]
%  \node (outer) [draw,fill=blue,fit=($(input) + (-0.95cm,4.5cm) (output)] {};%
%\end{scope}
%\node (ax)  [above of=input,label=above:$s\le 3$] {};

% draw plot for normal dist
\begin{axis}[
  no markers,
  domain=0:6,
  samples=100,
  ytick=\empty,
  xtick=\empty,
  xlabel={$(\mu,\sigma)$},
  width=3.5cm,
  at={($(input) + (-0.95cm,4.5cm)$)}
  ]
  \addplot [very thick,cyan!50!black, domain=0:6] {gauss(3,1)};

\end{axis}
    %\end{pgfonlayer}

%\end{scope}

% input -> latent
\draw[arr] (input) -- node[label=left:$q_\phi(z|x)$] {}($(input) + (0.0cm,4.0cm)$) {};

% latent -> output (reconstruction)
\draw[arr] ($(input) + (3.0cm,4.5cm)$) -- node[label=right:$p_\theta(x|z)$] {}(output) {};

%\end{scope}

\usetikzlibrary{calc}
\pgftransformreset
\node[inner sep=0pt,outer sep=0pt,minimum size=0pt,line width=0pt,text width=0pt,text height=0pt] at (current bounding box) {};
%add border to avoid cropping by pdflibnet
\foreach \border in {0.1}
  \useasboundingbox (current bounding box.south west)+(-\border,-\border) rectangle (current bounding box.north east)+(\border,\border);
\newwrite\metadatafile
\immediate\openout\metadatafile=\jobname_BB.txt
\path
  let
    \p1=(current bounding box.south west),
    \p2=(current bounding box.north east)
  in
  node[inner sep=0pt,outer sep=0pt,minimum size=0pt,line width=0pt,text width=0pt,text height=0pt,draw=white] at (current bounding box) {
\immediate\write\metadatafile{\p1,\p2}
};
\immediate\closeout\metadatafile
\end{tikzpicture}
\end{document}