### 2017-01-09 slides/figures/vae_with_rep.tex

\documentclass{standalone}

\usepackage{pgfplots}
\pgfplotsset{compat=1.14}
\usetikzlibrary{fit,backgrounds}

\usetikzlibrary{positioning, arrows.meta, calc, shapes.misc}

\begin{document}

\pgfmathdeclarefunction{gauss}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}

\tikzstyle{gb} = [draw,fill=blue!10,outer sep=0.1cm,minimum height=4em, minimum width=4em]
\tikzstyle{arr} = [thick,-{Latex[length=3mm,width=2mm]}]
\tikzstyle{cloud} = [draw, circle, node distance=2.5cm, minimum height=2em, gb]

\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\begin{tikzpicture}

%\begin{scope}[xshift=5cm,on grid]

% input space
\node[gb] (input) at (0cm,0cm) {\Huge$x$};
%\node[fill=black!20] (lbl_input) at (0,-1) {Input};

    %\begin{pgfonlayer}{background}
\node[above=3cm of input] (lat_start) {};
\node[above right=1.25cm and 0.85cm of lat_start] (lat_dist) {};
%\node[draw,right=3cm of lat_start] (lat_end) {};

% latent space
\node[] (lbl_lat) at ($(lat_dist) + (8.5cm,0.0cm)$) {\huge $z = \mu + \epsilon \odot \sigma$};

\node[above=of lbl_lat] (lbl_eps) {\huge$\epsilon \sim \mathcal{N}(0, I)$};

% output / reconstruction
\node[gb] (output) at ($(lbl_lat) + (0.0cm,-5.43cm)$) {\huge$\tilde x$};
%\node[fill=black!20] (lbl_rec) at ($(output) + (0,-1cm)$) {Reconstruction};

% draw plot for normal dist
\begin{axis}[
  no markers,
  domain=0:6,
  samples=100,
  ytick=\empty,
  width=3.5cm,
  xtick=\empty,
  xlabel={\huge$(\mu,\sigma)$},
  at={($(lat_start) + (-0.95cm,0.8cm)$)}
  ]
  \addplot [thick,cyan!50!black, domain=0:6] {gauss(3,1)};
  %\node[] (left_bot) at (axis cs:0.0,-0.15) {};
  %\node[] (left_top_aux) at (axis cs:0.0,0.2) {};
\end{axis}

  %\node [above right=of left_top_aux] (left_top) {Latent space};

    %\end{pgfonlayer}
%\begin{scope}[on background layer]
%  \node (outer) [dashed,draw,fill=blue!10,fit=(left_bot) (left_top) (lbl_lat)] {};
%\end{scope}

% input -> latent
\draw[arr] (input) -- node[label=left:\huge$q_\phi(z|x)$] {}(lat_start) {};

% latent -> latent
\draw[arr] (lat_dist) -- (lbl_lat) {};

\draw[arr] (lbl_eps) -- (lbl_lat);

%\node [label={[xshift=1.0cm, yshift=0.3cm]Label}] {Node};
% latent -> output (reconstruction)
\draw[arr] (lbl_lat) -- node[label=right:\huge$p_\theta(x|z)$] {}(output) {};

%\end{scope}

%\draw[dashed] (4.5,7.5) -- (4.5,-2);
\end{tikzpicture}
\end{document}