### 2017-01-20 slides/figures/latent_model.tex
updated & added figures

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta,fit}
\begin{document}
\tikzstyle{hidden} = [draw,circle,outer sep=0.05cm,inner sep=0.3cm,line width=1, thick, draw=black!55]
\tikzstyle{observed} = [draw,circle,outer sep=0.05cm,inner sep=0.3cm,line width=1, thick, draw=black!55, fill=black!20]
\tikzstyle{arr} = [thick,-{Latex[length=5mm,width=3mm]}]

\begin{tikzpicture}
\node[observed] (node_observed) {\Huge $x$};
\node[hidden,above=1.5cm of node_observed] (node_hidden) {\Huge $z$};

\draw[arr] (node_hidden) -- (node_observed) {};

\node[fit=(node_hidden)(node_observed), draw, rounded corners] {};
\end{tikzpicture}
\end{document}