### 2017-02-03 paper/chapters/vae/draw.tex
remove dumb emphasis

The Deep Recurrent Attention Writer\cite{draw:2015} (DRAW) is an advanced extension of the variational autoencoder.
It uses recurrent neural networks as encoder as well as decoder and includes a selective attention mechanism.
These advancements allow the model to observe an image through a series of \emph{glimpses} and draw to an output region determined by the attention mechanism.
Without going into much detail, this allows the model to sequentially draw images in a way which resembles much of human-like drawing or sketching.

%It includes recurrent neural networks and a selective attention mechanism to

%uses the selective attention mechanism in addition to a VAE with both recurrent encoder and recurrent decoder.
%DRAW allows to basically the pixels of an image sequentially very human-like.