### 2017-03-03 paper/chapters/vae/iwae.tex

% write about multiple stochastic layers

% write about k passes (just averages)

% write about IWAE, extends these k passes with importance weights
The Importance Weighted Autoencoder\cite{iwae:2015} (IWAE) extends on the VAE framework using $k$-sampled importance weighting of the log-likelihood.
% add intuition for IWAE, ruslan explained it very well during his DL summer school talk! (talk_Montreal_2016_Salakhutdinov.pdf)
Using importance weighting can be informally interpreted as putting more weight on data with high likelihood.

Improving on the theoretical results in the VAE framework, using IWAE with more samples will tighten the lower variational bound and therefore improve the overall performance of the model.
\begin{equation*}
\log p(x) \geq \mathcal{L}_{k+1}(x) \geq \mathcal{L}_{k}(x)
\end{equation*}

Additionally, if $\frac{p(h,x)}{q(h|x)}$ is bounded, then the model is guaranteed to reach the optimum as $k$ approaches infinity in equation \ref{eq:iwae}
\begin{equation}
\label{eq:iwae}
\lim_{k \rightarrow \infty} \mathcal{L}_{k}(x) = \log p(x)
\end{equation}

However, using more samples will result in higher computational cost and thus should be carefully evaluated.

%\vspace{5cm}